---
title: "Mission President Performance"
author: "Trenton Pulsipher"
date: "`r lubridate::today()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(trelliscopejs)
library(rbokeh)

raw <- read_csv("~/Documents/Development/R/data/missionaryInterview/KI_17SEP2018.csv") %>%
  as.tibble()
```


```{r areaAvgCalc, error = F, message = F, warning = F, echo = F}

# fix the Mission J outliers #
jBad1 <- raw[raw$mission_name == "J" & raw$week_mon_sun_end_date == "12/8/2011", "ni_reported"]
jBad2 <- raw[raw$mission_name == "J" & raw$week_mon_sun_end_date == "12/8/2011", "sa_reported"]

raw[raw$mission_name == "Q" & raw$week_mon_sun_end_date == "2/9/2012", "bc_reported"] <- 2
raw[raw$mission_name == "K" & raw$week_mon_sun_end_date == "2/11/2010", "ni_reported"] <- 113
raw[raw$mission_name == "J" & raw$week_mon_sun_end_date == "12/8/2011", "ni_reported"] <- jBad2
raw[raw$mission_name == "J" & raw$week_mon_sun_end_date == "12/8/2011", "sa_reported"] <- jBad1
raw[raw$mission_name == "L" & raw$week_mon_sun_end_date == "7/19/2007", "sa_reported"] <- 149    

areaAvg <- raw %>% 
  mutate(Date = mdy(week_mon_sun_end_date),
        `Young Missionaries` = elders + sisters,
        `Total Missionaries` = elders + sisters + seniors) %>%
  group_by(mission_name, mission_president, Date) %>%
  mutate(`Percent of Goal - New Investigators` = ni_reported / ni_goal,
         `Percent of Goal - Sacrament Attendance` = sa_reported / sa_goal,
         `Percent of Goal - Baptism Date Set` = bd_reported / bd_goal,
         `Percent of Goal - Baptism/Confirmation` = bc_reported / bc_goal,
         `New Investigators per Missionary` = ni_reported / `Total Missionaries`,
         `Sacrament Attendance per Missionary` = sa_reported / `Total Missionaries`,
         `Baptism Date Set per Missionary` = bd_reported / `Total Missionaries`,
         `Baptism/Confirmation per Missionary` = bc_reported / `Total Missionaries`) %>%
  rename(`New Investigators` = ni_reported,
         `Sacrament Attendance` = sa_reported,
         `Baptism Date Set` = bd_reported,
         `Baptism/Confirmation` = bc_reported,
         `New Investigators (Goal)` = ni_goal,
         `Sacrament Attendance (Goal)` = sa_goal,
         `Baptism Date Set (Goal)` = bd_goal,
         `Baptism/Confirmation (Goal)` = bc_goal,
         Elders = elders,
         Sisters = sisters,
         Seniors = seniors) %>%
  group_by(Date) %>%
  select(-mission_name, -mission_president, -week_mon_sun_end_date) %>%
  gather("metric", "value", 
         `Elders`, `Sisters`, `Seniors`, `Young Missionaries`, `Total Missionaries`,
         `New Investigators`, `Sacrament Attendance`, `Baptism Date Set`, `Baptism/Confirmation`,
         `New Investigators per Missionary`, `Sacrament Attendance per Missionary`, 
         `Baptism Date Set per Missionary`, `Baptism/Confirmation per Missionary`,
         `New Investigators (Goal)`, `Sacrament Attendance (Goal)`, `Baptism Date Set (Goal)`, `Baptism/Confirmation (Goal)`,
         `Percent of Goal - New Investigators`, `Percent of Goal - Sacrament Attendance`,
         `Percent of Goal - Baptism Date Set`, `Percent of Goal - Baptism/Confirmation`) %>%
  group_by(Date, metric) %>%
  summarise(AreaAvg = mean(value, na.rm = T))

```

Key Performance Indicators (KPIs) for each mission over time are shown below. Mission presidents are represented as a different color and hovering over each week's KPI point in the plot shows detail regarding the data. Mission health monitoring can help identify when a mission has the chills (KPIs too low compared to historical averages or area average) or a fever (KPIs too high prompting a different kind of concern and opportunity for training). KPIs are listed below the plots and summary for convenience.


```{r missionPresEffect, error = F, message = F, warning = F, echo = F, fig.height = 6, fig.width = 9}

byMetric <- raw %>% 
  mutate(Date = mdy(week_mon_sun_end_date),
        `Young Missionaries` = elders + sisters,
        `Total Missionaries` = elders + sisters + seniors) %>%
  group_by(mission_name, mission_president, Date) %>%
  mutate(`Percent of Goal - New Investigators` = ni_reported / ni_goal,
         `Percent of Goal - Sacrament Attendance` = sa_reported / sa_goal,
         `Percent of Goal - Baptism Date Set` = bd_reported / bd_goal,
         `Percent of Goal - Baptism/Confirmation` = bc_reported / bc_goal,
         `New Investigators per Missionary` = ni_reported / `Total Missionaries`,
         `Sacrament Attendance per Missionary` = sa_reported / `Total Missionaries`,
         `Baptism Date Set per Missionary` = bd_reported / `Total Missionaries`,
         `Baptism/Confirmation per Missionary` = bc_reported / `Total Missionaries`) %>%
  rename(`New Investigators` = ni_reported,
         `Sacrament Attendance` = sa_reported,
         `Baptism Date Set` = bd_reported,
         `Baptism/Confirmation` = bc_reported,
         `New Investigators (Goal)` = ni_goal,
         `Sacrament Attendance (Goal)` = sa_goal,
         `Baptism Date Set (Goal)` = bd_goal,
         `Baptism/Confirmation (Goal)` = bc_goal,
         Elders = elders,
         Sisters = sisters,
         Seniors = seniors) %>%
  gather("metric", "value", 
         `Elders`, `Sisters`, `Seniors`, `Young Missionaries`, `Total Missionaries`,
         `New Investigators`, `Sacrament Attendance`, `Baptism Date Set`, `Baptism/Confirmation`,
         `New Investigators per Missionary`, `Sacrament Attendance per Missionary`, 
         `Baptism Date Set per Missionary`, `Baptism/Confirmation per Missionary`,
         `New Investigators (Goal)`, `Sacrament Attendance (Goal)`, `Baptism Date Set (Goal)`, `Baptism/Confirmation (Goal)`,
         `Percent of Goal - New Investigators`, `Percent of Goal - Sacrament Attendance`,
         `Percent of Goal - Baptism Date Set`, `Percent of Goal - Baptism/Confirmation`)

byMetric %>%
  left_join(areaAvg %>%
              filter(is.finite(AreaAvg)), 
            by = c("Date", "metric")) %>%
  group_by(mission_name, metric) %>%
#  filter(mission_name %in% c("A", "B")) %>%
  rename(Pres = mission_president) %>%
  nest() %>%
  mutate(
    summaryStats = map_cog(data, ~ data_frame(
      mean = cog(mean(.x$value, na.rm = T), desc = "weekly avg"),
      median = cog(mean(.x$value, na.rm = T), desc = "weekly median"),
      stdDev = cog(sd(.x$value, na.rm = T), desc = "weekly standard deviation"),
      cv = cog(sd(.x$value, na.rm = T) / mean(.x$value, na.rm = T), desc = "weekly cv or sd/mean"))),
    panel = map_plot(data,
                     ~ figure(xlab = "", ylab = "") %>%
                       ly_points(x = Date, y = value, color = Pres, #mission_president, 
                                 hover = list(`Mission President` = Pres, #mission_president,
                                              `Date` = Date,
                                              metric = value),
                                 data = .x,
                                 legend = F) %>% 
                       ly_lines(x = Date, y = AreaAvg,
                                hover = list(`Date` = Date, 
                                            `Area Avg` = AreaAvg),
                                data = .x, 
                                legend = F))) %>%
  trelliscope(name = "KPIs by Mission", self_contained = T)

```

### Summary of Findings

Ideally something like a summary table auto-generated to indicate mission health would be shared internally and possibly with area presidencies regularly. Below are some of my quick observations that give a flavor as to what could be provided in such a table. Comments here reflect only a portion of the missions and only a few KPIs.

- Mission A: only baptisms appear near the area average, mission president effect is very low
- Mission B: all KPIs are generally lower than area average, strong mission president effect, recent improvements made by current mission president
- Mission C: high variability in new investigators and baptism date set KPIs, generally at or below area average, mission president effect observed on all KPIs except baptisms
- Mission H: new investigators KPI is very positive, generally at or above area average on all KPIs, weak mission president effect
- Mission I: strong mission president effect observed, KPIs appear above area average, recent correction of KPIs observed, overall a very interesting and somewhat unique KPI profile
- Mission K: KPIs appear at or above area average, baptisms seem consistent but the other KPIs show a strong mission president effect

This was not an exhaustive analysis and should have been developed into a more comprehensive and cohesive table of results. Ideally those results would like directly to a visualization of the summarized content.


#### KPIs

KPIs available for viewing here include:

* New Investigators
* Sacrament Attendance
* Baptism Date Set
* Baptism/Confirmation

The four KPIs adjusted for the number of missionaries:

* New Investigators per Missionary
* Sacrament Attendance per Missionary
* Baptism Date Set per Missionary
* Baptism/Confirmation per Missionary

Weekly mission goals:

* New Investigators (Goal)
* Sacrament Attendance (Goal)
* Baptism Date Set (Goal)
* Baptism/Confirmation (Goal)

Percent of mission goal attained:

* Percent of Goal - New Investigators
* Percent of Goal - Sacrament Attendance
* Percent of Goal - Baptism Date Set
* Percent of Goal - Baptism/Confirmation`

Number of missionaries per mission:

* Elders
* Sisters
* Seniors
* Young Missionaries (Elders + Sisters)
* Total Missionaries (Elders + Sisters + Seniors)

Conversion Efficiency (not included in this analysis): 

* New Investigators / Baptism/Confirmation
* Sacrament Attendance / Baptism/Confirmation
* Baptism Date Set / Baptism/Confirmation
