---
title: "Notes"
author: "Trenton Pulsipher"
date: "9/20/2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)

raw <- read_csv("~/Documents/Development/R/data/missionaryInterview/KI_17SEP2018.csv") %>%
  as.tibble()
```

## Observations/Thoughts

#### Mission President Effect


###### New Investigators

Clearly a mission president effect is observed, note the color changes and regression lines by mission president for each mission.

```{r missionPresEffect1, error = F, message = F, warning = F, echo = F, fig.height = 6, fig.width = 9}
raw %>% 
   mutate(Date = mdy(week_mon_sun_end_date),
          YoungMissionaries = elders + sisters,
          TotalMissionaries = elders + sisters + seniors) %>%
   ggplot(aes(x = Date, y = ni_reported, colour = mission_president)) +
     geom_point(alpha = .25) +
     geom_smooth(method = "lm", se = F) +
     theme_bw() +
     facet_wrap(~ mission_name, scales = "free_y") +
     theme(legend.position = "none") +
     labs(x = "", 
          y = "New Investigators Reported", 
          title = "Weekly Number of New Investigators by Mission")
```

However we should confirm that the mission president effect isn't overly influenced/confounded by the number of missionaries. The plot below confirms that the mission president has a potentially strong effect on the number of new investigators within a mission.

```{r missionPresEffect2, error = F, message = F, warning = F, echo = F, fig.height = 6, fig.width = 9}
raw %>% 
   mutate(Date = mdy(week_mon_sun_end_date),
          YoungMissionaries = elders + sisters,
          TotalMissionaries = elders + sisters + seniors, 
          ni_reported_numMiss = ni_reported / TotalMissionaries) %>%
   ggplot(aes(x = Date, y = ni_reported_numMiss, colour = mission_president)) +
     geom_point(alpha = .25) +
    geom_smooth(method = "lm", se = F) +
     theme_bw() +
     facet_wrap(~ mission_name, scales = "free_y") +
     theme(legend.position = "none") +
     labs(x = "", 
          y = "New Investigators Reported per Missionary", 
          title = "Weekly Number of New Investigators per Missionary by Mission")
```


###### Baptisms/Confirmations

Again a mission president effect is observed.

```{r missionPresEffect3, error = F, message = F, warning = F, echo = F, fig.height = 6, fig.width = 9}
raw %>% 
   mutate(Date = mdy(week_mon_sun_end_date),
          YoungMissionaries = elders + sisters,
          TotalMissionaries = elders + sisters + seniors, 
          bc_reported_numMiss = bc_reported / TotalMissionaries) %>%
   ggplot(aes(x = Date, y = bc_reported_numMiss, colour = mission_president)) +
     geom_point(alpha = .25) +
    geom_smooth(method = "lm", se = F) +
     theme_bw() +
     facet_wrap(~ mission_name, scales = "free_y") +
     theme(legend.position = "none") +
     labs(x = "", 
          y = "Baptisms/Confirmations Reported per Missionary", 
          title = "Weekly Number of Baptisms/Confirmations per Missionary by Mission")
```


#### Data Cleaning/Correcting Opportunities

###### Potential Outliers 

Outliers are possible and we would want to be careful about their effect on predictions/goals. NOTE THE OUTLIERS HERE

```{r outliers1, error = F, message = F, warning = F, echo = F, fig.height = 6, fig.width = 9}
raw %>% 
  mutate(Date = mdy(week_mon_sun_end_date),
        YoungMissionaries = elders + sisters,
        TotalMissionaries = elders + sisters + seniors) %>%
  group_by(mission_name, mission_president, Date) %>%
  select(contains("reported")) %>%
  rename(`New Investigators` = ni_reported,
         `Sacrament Attendance` = sa_reported,
         `Baptism Date Set` = bd_reported,
         `Baptism/Confirmation` = bc_reported) %>%
  gather("metric", "value", `New Investigators`, `Sacrament Attendance`, `Baptism Date Set`, `Baptism/Confirmation`) %>%
  ggplot(aes(x = Date, y = value, group = mission_name)) +
    geom_line() +
    geom_smooth(method = "loess") +
    theme_bw() +
    facet_grid(metric ~ mission_name, scales = "free_y") +
    theme(legend.position = "none") +
    labs(x = "", y = "")
```



Quote from *Preach My Gospel* about goals and how that influenced my suggestions regarding mission goals

Incorporate a machine learning model or time-series model for prediction.
